---
title: "pa 4"
author: "Ying Zhang"
institute: "Rutgers University"
date: "`r Sys.Date()`"
output:
  xaringan::moon_reader:
    lib_dir: libs
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---
# 1. Setup
```{r, message=FALSE}

library("tidyverse")
library("ds4ling")
library("untidydata")
library("here")
library("broom")
data <- read.csv(here("data_raw", "fricatives","fricatives.csv"))
```
---
# 2. Tidy the data
.tiny[
```{r}
data_tidy <- data |> 
  pivot_longer(cols = c("s_cog","s_skewness","sh_cog","sh_skewness"),
    names_to = "sounds", values_to = "values"
  ) |> 
  separate(col = sounds, into = c("phoneme", "measurement"), sep = "_"
  ) |> 
  pivot_wider(
    names_from = measurement,
    values_from = values
  ) |> 
  mutate_if(is.numeric,round, digits = 2)

write_csv(data_tidy, "/Users/yingzhang/Desktop/R/pa4/pa4/data_tidy/fricatives_tidy.csv")

head(data_tidy)
```
]
---
# 3. Descriptive statistics
```{r, fig.height=2, fig.width=4, out.width="50%"}
data_tidy |> 
  group_by(phoneme) |> 
  summarise(avg_cog = mean(cog),
            std_cog = sd(cog),
            avg_ske = mean(skewness),
            std_ske = mean(skewness)) |> 
  mutate_if(is.numeric, round, digits =2) |> 
  knitr::kable()
```

---
# 4. Center of gravity as a function of phoneme
```{r}
data_tidy |> 
ggplot() + aes(x = phoneme, y = cog) + geom_boxplot() + labs(x = "Phoneme", y = "Center of gravity")
```
---
# 5. Skewness as a function of phoneme
```{r}
data_tidy |> ggplot() + aes(x = phoneme, y = skewness) + stat_summary(fun = mean, geom = "point") + stat_summary(fun.data = mean_se, geom = "errorbar", width = 0.2)
```

---
# 6. Center of gravity as a function of skewness for the [s] segments
## Summary
```{r}
data_s <- data_tidy |>pivot_wider(names_from = phoneme, values_from = c(cog, skewness))

mod1 <- lm(cog_s ~ skewness_s, data = data_s)

mod_summary <- tidy(mod1) 

  knitr::kable(mod_summary, digits = 3 )
```
---
# 6. Center of gravity as a function of skewness for the [s] segments 
## Scatterplot

```{r}
data_s |> 
  ggplot() +
  aes(x = skewness_s, y = cog_s) +
  geom_point() + geom_smooth(method = "lm") +
  labs(x = "Skewness", y = "Center of gravity") + 
  ggtitle("Center of gravity as a function of skewness for [s]")
```

---
# 7. Model diagnostics
```{r}
summary(mod1)
```

---
# 7. Model diagnostics

```{r}
ds4ling::diagnosis(mod1)
```

---
# 8. The results
The data of fricatives was analyzed using a linear model to examine the effect of skewness on center of gravity in terms of [s]. In the plot of residuals (the left panel), the residuals are scattered randomly with no clear pattern, indicating a linear relationship between the skewness and center of gravity. The histogram has a bell-like shape centered close to 0, which suggests that the residuals are approximately normally distributed. The Q-Q plot shows that most points fall closely along the line only with minor deviation in the tail. It indicates that the residuals are highly normally distributed.

The summary shows that the model can account for 75.2% of the variance (Multiple R-squared: 0.752, Adjusted R-squared: 0,7468). Center of gravity has a linear relationship with skewness in terms of [s] (t = 12.064, p < .001). Every one unit change of skewness is associated with an increase of center of gravity by 1678.2 units. 
---
# 9. Scatterplot (assumptions)
```{r}
data_asm <- read.csv(here("data_raw", "fricatives","assumptions.csv"))

data_asm |> 
  ggplot() +
  aes(x = x, y = y) +
  geom_point()

```
---
# 9. Explanation

x and y are not linearly correlated as shown in the plot. One value of x can correspond to more than one value of y so there is no model that fits this data but a dinosaur can!
